3D Regression Heat Map Analysis of Population Study Data

### 2015-06-22 draft.tex
replaced occurences of 'cube'

\title{Regression Cube Analysis of Population Study Data}

@@

%other entries to be set up for journal
\shortauthortitle{Klemm \MakeLowercase{\textit{et al.}}: Regression Cube Analysis of Population Study Data}
%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}

@@

%\com{New Solution}.
%%
We propose an analysis approach of epidemiological data sets by incorporating all features in an exhaustive regression-based analysis.
%%
This approach combines all \emph{independent features} with respect to a \emph{target feature}.
%%
It provides a visualization that reveals insights into the data by highlighting relationships.
%%
The \emph{Regression Cube}, a novel 3D visual encoding, acts as an overview of the whole data set.
%%
The Cube shows all combinations of two to three independent features w.r.t. a specific target disease.
%%
%A 3D visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{Regression Cube}.
%%
Slicing through the \emph{Regression Cube} allows for the detailed analysis of the underlying relationships.
%%
Expert knowledge about disease-specific hypotheses can be included into the analysis by adjusting the regression model formulas.
%%
Furthermore, the influences of features can be assessed using a difference view comparing different calculation results.
%%
%\com{Validation \& Results}.
%%
We applied our \emph{Regression Cube} method to a hepatic steatosis data set to reproduce results from a data mining-driven analysis.
%%
A qualitative analysis was conducted on a breast fat data set.
%%
We were able to derive new hypotheses about relations between breast fat density and breast lesions with breast cancer.
%%
%\com{Implications}.
%%
With the \emph{Regression Cube}, we present a visual overview of epidemiological data that allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
} % end of abstract

@@

\design{
\begin{itemize}
	%%
  \item An overview visualization design based on feedback of epidemiological domain experts to support hypothesis-generation w.r.t. a target feature using regression models.
	%%
  \item Applying domain knowledge in the analysis using freely adjustable regression formulas.
	%%
  \item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details on demand to assess the model quality.
	%%
  \item An openly available web-based system, which incorporates open formats to be applicable for data of other application domains.
	%%
\end{itemize}
}
% \begin{itemize}
% 	%%
% 	\item An overview visualization technique representing feature interactions using target features.
% 	%%
% 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
% 	%%
% 	\item Freely adjustable regression formulas to provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
% 	%%
% 	\item The analysis of confounding features by providing comparison views between different formula results.
% 	%%
% 	\item The expansion and adaption of our method to any data with our open web-based system.
% 	%%
% \end{itemize}
%%
\noindent The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
%%
They extend 2D scatter plots to display detailed information about a regression model.
%%
In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
%%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow and requirements.
%%
\subsection{Epidemiological Workflow} \label{sec:EpidemiologicalWorkflow}
%%
Epidemiological research is performed by experts from different academic disciplines, such as epidemiologists, physicians, statisticians, and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features with statistical methods.
%%
As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{enumerate}
	\item Clinicians/Epidemiologists derive hypotheses via clinical observations, experimental studies or literature research.
	%\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
	\item Statisticians assess the association of the derived features w.r.t the investigated disease.
\end{enumerate}
%%
Relative risks can be determined if a statistical resilient association of features with a condition is extracted.
%%
They indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow.
%%
Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

@@

%% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
Similarly, Piringer et al.~\cite{Piringer} proposed methods for visualizing regression analysis results and properties for developing car engines.
%%
Their main goal is to assess the pairwise influence of independent features w.r.t. the target feature using a plot matrix displaying models as contours.
%%
%They also incorporate 3D visualizations for each pairwise combination, but mainly because of its popularity with the target domain engineers.
%%
Linked views of model deviations allow to select outliers.
%%
This limits the method to comparing a few models at once as the plot matrix gets complex with increasing feature number.
%%
% The main difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%
% We analyze a large amount of simpler models in terms of \emph{different} features.
% %%
% They also focus on metric features, while we process categorical data as well.
%%
\magenta{
%%
Guo et al.~\cite{Guo} presented a multi-space visualizations to find linear relationships in the data with focus on extracting groups of best fit.
%%
The data space is visualized using a scatter plot matrix.
%%
Linear models are calculated by defining dependent and independent features.
%%
The model view allows for assessing different models by color coding distances to the line-of-fit.
%%
The model-parameters can then be fine-tuned using line graphs, histograms and model projections.
%%
}
%% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
We use the term `Regression Cube' for a comparative 3D representation of multiple regression models.
%%
Chan et al.~\cite{Chan} use the same term to describe an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
%%
They group subjects using a set of interaction techniques as well as clustering algorithms to calculate sub-groups, which can then be compared using their cube representation.
%%
Similar to Piringer et al.~\cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
%%
Insight is derived by subject grouping, which spawns new cube correlations and therefore allows drilling down to the data.
%%
%In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
%%
%We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
%%
Piringer et al.~\cite{Piringer} Guo et al.~\cite{Guo} and Chan et al.~\cite{Chan} focus on finding and tuning a model for a specific relationship, while we search for models w.r.t. all combinations in the data set.
%%
Instead of analyzing one complex model in detail, we process a large amount of models in terms of \emph{different} features.
%We difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%

@@

\section{Regression Cube Analysis of Population Study Data}

@@

The \emph{Regression Cube} is designed to provide overview visualizations to support hypothesis generation.

@@

% This section covers the design of the Regression Cube as well as how epidemiological requirements and  domain expert input was utilized to influence design decisions.

@@

The \emph{Regression Cube} design was developed iteratively based on feedback of epidemiologists by using the prototype in joint analysis sessions on their data sets.

@@

%The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.

@@

As described in Section~\ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing relationships in epidemiological data.
%%
A regression model is based on expert knowledge; there is no singular rule how to apply models to a given set of features. Thus, they have to be applied with care.
%%
% \design{
% %%
% Key requirements based on the epidemiological application domain on the Regression Cube are:
% %%
% \begin{itemize}
% 	\item Show relationships of features w.r.t. their explanatory power towards the target feature (e.g., a disease or unhealthy condition).
% 	\item Allow for modeling assumptions regarding the target feature.
% 	\item The tool has to be easily accessible and simultaneously secure the data its processing.
% \end{itemize}
% %%
% }
\subsection{Cube Description Using Regression Formula Notation}
%%
\design{The key requirement of modeling expert knowledge} is integrated into the analysis using \emph{regression formulas}.
%Expert knowledge is integrated into the analysis using regression formulas.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
%%

@@

Identifying irrelevant features and excluding them from the feature space considerably reduces computational costs and yields a comprehensible \emph{Regression Cube} representation.

@@

%\subsection{Abstracting Regression Results Using $R^2$}
\subsection{Abstracting Regression Results}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/cube}
 \caption{
 (a) Overview visualization using a 2D heat map for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The descriptive metrics extracted from the regression model are mapped to color saturation (for $R^2$, saturated color shows a strong correlation).
 %%
 (b) Now, $Z$ is set to all features $n$ and yields $n$ 2D heat map visualizations.
 %%
 These represent the slices in our cube.
 %%
 The metric describing the regression model of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Cube}
\end{figure}
%%
\noindent The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%%
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a 2D heat map.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model.
%%
This allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ \add{, adjusted $R^2$ and AIC} value are metrics allowing for this kind of assessment.
%%
% \add{
% %%
% Users with in-depth knowledge about regression models are given the option to render the \emph{Regression Cube} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
% %%
% Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
% %%
% To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
% }
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
%%
We can apply the same 2D heat map by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Cube}a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%
\design{
%%
Based on expert feedback on early versions of this view, the amount of features used to compare regression models was extended.
%%
Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
%%
}
%%
\add{
%%
$Adjusted$ $R^2$ can be represented in the same way, since they share the scale.
%%
$AIC$ values are not restricted to a normalized scale.
%%
They have to be normalized in order to map it onto color saturation.
%%
%The resulting scale may be distort by outliers, which requires facilities for allowing the user to select scale ranges.
The resulting scale may be distorted by outliers derived from poor regression models.
%%
To tackle this problem, we provide a slider input, which maps the transfer function of the metric to color saturation based on user-selected ranges.
%%
Outliers can be cut off to emphasize ranges of interest.
%%
Small $AIC$ values indicate a good model.
%%
Hence, we inverted the transfer function mapping it to color, mapping low $AIC$ features to saturated colors.
%%
}
\design{
%%
To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
%%
}

\paragraph{3D (Cube) View.}
%%
Introducing $Z$ creates a 3D \emph{Regression Cube} (Fig.~\ref{fig:Cube}b).
%%
\add{The selected metric (per default set to $R^2$)} of each cube entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
%%
\magenta{Object size is not used to encode information because it would result in a cluttered view.}
%%
\design{
%%
Epidemiologists with background on statistics argued, that the visualization of descriptive metrics derived from different regression methods (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
Therefore, we decided to map metrics of different regression methods on distinct colors (i.e., orange for linear regression and blue for logistic regression).
%%
}
%%
%The visualization of descriptive metrics derived from different regression cubes (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
%Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., orange for linear regression and blue for logistic regression).
%%
Thus, the cube can be easily extended using other regression types.
%%
For cubes with a fixed target feature, e.g., $Cancer \sim X + Y + Z$, no such encodings are required and the $z$ dimension can be compared directly.
%%
\magenta{
As mentioned previously, the dimension reduction using the CFS algorithm potentially removes features, which are of interest for the domain expert.
%%
To tackle this problem, the $z$ dimension of the cube contains all features of the data set, allowing to assess their influence on the models.
%%
The $x$ and $y$ dimensions are restricted to the features extracted from the CFS algorithm.
}
%%
\\\\
%%
Our goal is to create an overview visualization for a data set.
%%
We also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
%%
These two approaches do not exclude each other, they rather underline the difference in purpose of the chosen formula.
%%
The different analysis approaches require different starting points using the \emph{Regression Cube}.

\subsection{Analysis Workflow} \label{sec:Workflow}
%%
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/workflow}
 \caption{
 Different workflow types of the analysis using \emph{Regression Cubes}.
 %%
 [\textbf{1.AF}] The workflow starts by declaring a formula to specify a hypothesis, or to use a predefined formula for a hypothesis-free analysis.
 %%
 [\textbf{2.SI}] The \emph{Regression Cube} is then visualized.
 %%
 The user has the option to either adjust the formula or to derive details on demand on a specific regression.
 %%
 [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or a selection of a difference view.
 %%
 The latter is used to compare regression cubes.
 %%
 [\textbf{4.DD}] Details about features using the 2D heat map representation yield insights and hypotheses about feature relations.
 }
  \label{fig:Workflow}
\end{figure}
%%
\noindent Our \emph{Regression Cube} model is well suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:

\textbf{Analyze First} [\textbf{1.AF}]. Choosing an initial regression formula triggers the \emph{Regression Cube} calculation on the given data set, filtering the dimensions of the dependent feature through the CFS algorithm.

\textbf{Show the Important} [\textbf{2.SI}]. The 3D cube visualization acts as an overview over the whole data set.
%%
Here, regression models with large regression metric values can be spotted fast, steering the user's attention to the respective slice.

@@

\textbf{Details on Demand} [\textbf{4.DD}]. Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover modal on a currently selected data point).
\\\\
We use the squared bracket abbreviation for each step to denote the affiliation to the system design section later on.
%%
As shown in Fig.~\ref{fig:Workflow}, the workflow is highly iterative.
%%
Observations in the 2D heat map or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or to create a new \emph{Regression Cube} for a difference view.
%%
\paragraph{Hypothesis-Free and Hypothesis-Based Analysis.}
%%
\design{
%%
Early analysis sessions yielded two approaches on analyzing the data.
%%
The classic approach is \emph{hypothesis-based}, where the expert already knows the data and potential associations (e.g. reproducing knowledge about hepatic steatosis risk factors based on known risk factors).
%%
The \emph{hypothesis-free} analysis is well suited for deriving new insight into the data, such as identifying confounding features or potential targets (e.g. deriving risk factors for breast cancer associated features).
%%
}
%%
%Input formulas reflect \emph{hypotheses} about the data.
\emph{Hypotheses} about the data are reflected using input formulas.
%%
Using the operators, dynamic variables and dataset features, many different assumptions can be expressed.
%%
To support the \emph{hypothesis-free} analysis, we provide a default formula:

$Z \sim X + Y$.
%%
This cube represents all possible combinations of two independent features w.r.t. all features in the data set, since we do not know which features are of interest.
%%
Each slice represents a different target feature.
%%
It is therefore suitable for an explorative analysis to give a general impression about relationships in the data set.

@@

%%
$Cancer \sim X + Y + Z$ is the formulation of a hypothesis, where the specific feature \textit{Cancer} is analyzed.
%%
All combinations of three independent features with the target are analyzed through the cube.

$Cancer \sim X + Y + Z + feature_1:feature_2$ encodes more assumptions.
%%
This formula models the hypothesis of an interaction between $feature_1$ and $feature_2$ (denoted with `$:$') being relevant for the target feature, but it is not clear how other feature combinations influence the result.
%%
Therefore, the cube incorporates this interaction for all $X$, $Y$ and $Z$ values as independent features.
%%

$Cancer \sim X + Y + Z$ subtracted with the regression metric from $Cancer \sim Age$ excludes the confounding effect that age has in view of the target $Cancer$ feature.
%%
This is achieved through cube comparison.
%%

>>>>

\title{3D Regression Heat Map Analysis of Population Study Data}

@@

%other entries to be set up for journal
\shortauthortitle{Klemm \MakeLowercase{\textit{et al.}}: 3D Regression Heat Map Analysis of Population Study Data}
%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}

@@

%\com{New Solution}.
%%
We propose an analysis approach of epidemiological data sets by incorporating all features in an exhaustive regression-based analysis.
%%
This approach combines all \emph{independent features} with respect to a \emph{target feature}.
%%
It provides a visualization that reveals insights into the data by highlighting relationships.
%%
The \emph{3D Regression Heat Map}, a novel 3D visual encoding, acts as an overview of the whole data set.
%%
It shows all combinations of two to three independent features w.r.t. a specific target disease.
%%
%A 3D visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{3D Regression Heat Map}.
%%
Slicing through the \emph{3D Regression Heat Map} allows for the detailed analysis of the underlying relationships.
%%
Expert knowledge about disease-specific hypotheses can be included into the analysis by adjusting the regression model formulas.
%%
Furthermore, the influences of features can be assessed using a difference view comparing different calculation results.
%%
%\com{Validation \& Results}.
%%
We applied our \emph{3D Regression Heat Map} method to a hepatic steatosis data set to reproduce results from a data mining-driven analysis.
%%
A qualitative analysis was conducted on a breast fat data set.
%%
We were able to derive new hypotheses about relations between breast fat density and breast lesions with breast cancer.
%%
%\com{Implications}.
%%
With the \emph{3D Regression Heat Map}, we present a visual overview of epidemiological data that allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
} % end of abstract

@@

\design{
\begin{itemize}
	%%
  \item An overview visualization design based on feedback of epidemiological domain experts to support hypothesis-generation w.r.t. a target feature using regression models.
	%%
  \item Applying domain knowledge in the analysis using freely adjustable regression formulas.
	%%
  \item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details on demand to assess the model quality.
	%%
  \item An openly available web-based system, which incorporates open formats to be applicable for data of other application domains.
	%%
\end{itemize}
}
% \begin{itemize}
% 	%%
% 	\item An overview visualization technique representing feature interactions using target features.
% 	%%
% 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
% 	%%
% 	\item Freely adjustable regression formulas to provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
% 	%%
% 	\item The analysis of confounding features by providing comparison views between different formula results.
% 	%%
% 	\item The expansion and adaption of our method to any data with our open web-based system.
% 	%%
% \end{itemize}
%%
% \noindent The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
% %%
% They extend 2D scatter plots to display detailed information about a regression model.
% %%
% In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
% %%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow and requirements.
%%
\subsection{Epidemiological Workflow} \label{sec:EpidemiologicalWorkflow}
%%
Epidemiological research is performed by experts from different academic disciplines, such as epidemiologists, physicians, statisticians, and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features with statistical methods.
%%
As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{enumerate}
	\item Clinicians/Epidemiologists derive hypotheses via clinical observations, experimental studies or literature research.
	%\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
	\item Statisticians assess the association of the derived features w.r.t the investigated disease.
\end{enumerate}
%%
Relative risks can be determined if a statistical resilient association of features with a condition is extracted.
%%
They indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow.
%%
Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

@@

%% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
Similarly, Piringer et al.~\cite{Piringer} proposed methods for visualizing regression analysis results and properties for developing car engines.
%%
Their main goal is to assess the pairwise influence of independent features w.r.t. the target feature using a plot matrix displaying models as contours.
%%
%They also incorporate 3D visualizations for each pairwise combination, but mainly because of its popularity with the target domain engineers.
%%
Linked views of model deviations allow to select outliers.
%%
This limits the method to comparing a few models at once as the plot matrix gets complex with increasing feature number.
%%
% The main difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%
% We analyze a large amount of simpler models in terms of \emph{different} features.
% %%
% They also focus on metric features, while we process categorical data as well.
%%
\magenta{
%%
Guo et al.~\cite{Guo} presented a multi-space visualizations to find linear relationships in the data with focus on extracting groups of best fit.
%%
The data space is visualized using a scatter plot matrix.
%%
Linear models are calculated by defining dependent and independent features.
%%
The model view allows for assessing different models by color coding distances to the line-of-fit.
%%
The model-parameters can then be fine-tuned using line graphs, histograms and model projections.
%%
}
%% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
%We use the term `Regression Cube' for a comparative 3D representation of multiple regression models.
%%
Chan et al.~\cite{Chan} propose the `Regression Cube', an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
%%
They group subjects using a set of interaction techniques as well as clustering algorithms to calculate sub-groups, which can then be compared using their cube representation.
%%
Similar to Piringer et al.~\cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
%%
Insight is derived by subject grouping, which spawns new cube correlations and therefore allows drilling down to the data.
%%
%In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
%%
%We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
%%
Piringer et al.~\cite{Piringer} Guo et al.~\cite{Guo} and Chan et al.~\cite{Chan} focus on finding and tuning a model for a specific relationship, while we search for models w.r.t. all combinations in the data set.
%%
Instead of analyzing one complex model in detail, we process a large amount of models in terms of \emph{different} features.
%We difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%

@@

\section{3D Regression Heat Map Analysis of Population Study Data}

@@

The \emph{3D Regression Heat Map} is designed to provide overview visualizations to support hypothesis generation.

@@

% This section covers the design of the 3D Regression Heat Map as well as how epidemiological requirements and  domain expert input was utilized to influence design decisions.

@@

The \emph{3D Regression Heat Map} design was developed iteratively based on feedback of epidemiologists by using the prototype in joint analysis sessions on their data sets.

@@

%The basic idea of our \emph{3D Regression Heat Map} is to provide an overview visualization of large cohort study data sets towards target features.

@@

As described in Section~\ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing relationships in epidemiological data.
%%
A regression model is based on expert knowledge; there is no singular rule how to apply models to a given set of features. Thus, they have to be applied with care.
%%
% \design{
% %%
% Key requirements based on the epidemiological application domain on the 3D Regression Heat Map are:
% %%
% \begin{itemize}
% 	\item Show relationships of features w.r.t. their explanatory power towards the target feature (e.g., a disease or unhealthy condition).
% 	\item Allow for modeling assumptions regarding the target feature.
% 	\item The tool has to be easily accessible and simultaneously secure the data its processing.
% \end{itemize}
% %%
% }
\subsection{Regression Heat Map Description Using Regression Formula Notation}
%%
\design{The key requirement of modeling expert knowledge} is integrated into the analysis using \emph{regression formulas}.
%Expert knowledge is integrated into the analysis using regression formulas.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
%%

@@

Identifying irrelevant features and excluding them from the feature space considerably reduces computational costs and yields a comprehensible \emph{3D Regression Heat Map} representation.

@@

%\subsection{Abstracting Regression Results Using $R^2$}
\subsection{Abstracting Regression Results}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/cube}
 \caption{
 (a) Overview visualization using a 2D heat map for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The descriptive metrics extracted from the regression model are mapped to color saturation (for $R^2$, saturated color shows a strong correlation).
 %%
 (b) Now, $Z$ is set to all features $n$ and yields $n$ 2D heat map visualizations.
 %%
 These represent the slices in our \emph{3D Regression Heat Map}.
 %%
 The metric describing the regression model of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Cube}
\end{figure}
%%
\noindent The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%%
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a 2D heat map.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{3D Regression Heat Map} analysis we are interested in the quality-of-fit of the resulting model.
%%
This allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ \add{, adjusted $R^2$ and AIC} value are metrics allowing for this kind of assessment.
%%
% \add{
% %%
% Users with in-depth knowledge about regression models are given the option to render the \emph{3D Regression Heat Map} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
% %%
% Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
% %%
% To avoid shutting out users unfamiliar with these metrics, the \emph{3D Regression Heat Map} is set per default to show $R^2$ values.
% }
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
%%
We can apply the same 2D heat map by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Cube}a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%
\design{
%%
Based on expert feedback on early versions of this view, the amount of features used to compare regression models was extended.
%%
Therefore, these experts can investigate the heat map with emphasis on specific aspects of the model.
%%
}
%%
\add{
%%
$Adjusted$ $R^2$ can be represented in the same way, since they share the scale.
%%
$AIC$ values are not restricted to a normalized scale.
%%
They have to be normalized in order to map it onto color saturation.
%%
%The resulting scale may be distort by outliers, which requires facilities for allowing the user to select scale ranges.
The resulting scale may be distorted by outliers derived from poor regression models.
%%
To tackle this problem, we provide a slider input, which maps the transfer function of the metric to color saturation based on user-selected ranges.
%%
Outliers can be cut off to emphasize ranges of interest.
%%
Small $AIC$ values indicate a good model.
%%
Hence, we inverted the transfer function mapping it to color, mapping low $AIC$ features to saturated colors.
%%
}
\design{
%%
To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Heat Map} is set per default to show $R^2$ values.
%%
}

\paragraph{3D View.}
%%
Introducing $Z$ creates a 3D heat map (Fig.~\ref{fig:Cube}b).
%%
\add{The selected metric (per default set to $R^2$)} of each heat map entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
%%
\magenta{Object size is not used to encode information because it would result in a cluttered view.}
%%
\design{
%%
Epidemiologists with background on statistics argued, that the visualization of descriptive metrics derived from different regression methods (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
Therefore, we decided to map metrics of different regression methods on distinct colors (i.e., orange for linear regression and blue for logistic regression).
%%
}
%%
%The visualization of descriptive metrics derived from different 3D Regression Heat Maps (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
%Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., orange for linear regression and blue for logistic regression).
%%
Thus, the visualization can be easily extended using other regression types.
%%
For \emph{3D Regression Heat Maps} with a fixed target feature, e.g., $Cancer \sim X + Y + Z$, no such encodings are required and the $z$ dimension can be compared directly.
%%
\magenta{
As mentioned previously, the dimension reduction using the CFS algorithm potentially removes features, which are of interest for the domain expert.
%%
To tackle this problem, the $z$ dimension of the visualization contains all features of the data set, allowing to assess their influence on the models.
%%
The $x$ and $y$ dimensions are restricted to the features extracted from the CFS algorithm.
}
%%
\\\\
%%
Our goal is to create an overview visualization for a data set.
%%
We also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
%%
These two approaches do not exclude each other, they rather underline the difference in purpose of the chosen formula.
%%
The different analysis approaches require different starting points using the \emph{3D Regression Heat Map}.

\subsection{Analysis Workflow} \label{sec:Workflow}
%%
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/workflow}
 \caption{
 Different workflow types of the analysis using \emph{3D Regression Heat Maps}.
 %%
 [\textbf{1.AF}] The workflow starts by declaring a formula to specify a hypothesis, or to use a predefined formula for a hypothesis-free analysis.
 %%
 [\textbf{2.SI}] The \emph{3D Regression Heat Map} is then visualized.
 %%
 The user has the option to either adjust the formula or to derive details on demand on a specific regression.
 %%
 [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or a selection of a difference view.
 %%
 The latter is used to compare 3D Regression Heat Maps.
 %%
 [\textbf{4.DD}] Details about features using the 2D heat map representation yield insights and hypotheses about feature relations.
 }
  \label{fig:Workflow}
\end{figure}
%%
\noindent Our \emph{3D Regression Heat Map} model is well suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:

\textbf{Analyze First} [\textbf{1.AF}]. Choosing an initial regression formula triggers the \emph{3D Regression Heat Map} calculation on the given data set, filtering the dimensions of the dependent feature through the CFS algorithm.

\textbf{Show the Important} [\textbf{2.SI}]. The 3D visualization acts as an overview over the whole data set.
%%
Here, regression models with large regression metric values can be spotted fast, steering the user's attention to the respective slice.

@@

\textbf{Details on Demand} [\textbf{4.DD}]. Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover modal on a currently selected data point).
\\\\
We use the squared bracket abbreviation for each step to denote the affiliation to the system design section later on.
%%
As shown in Fig.~\ref{fig:Workflow}, the workflow is highly iterative.
%%
Observations in the 2D heat map or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or to create a new \emph{3D Regression Heat Map} for a difference view.
%%
\paragraph{Hypothesis-Free and Hypothesis-Based Analysis.}
%%
\design{
%%
Early analysis sessions yielded two approaches on analyzing the data.
%%
The classic approach is \emph{hypothesis-based}, where the expert already knows the data and potential associations (e.g. reproducing knowledge about hepatic steatosis risk factors based on known risk factors).
%%
The \emph{hypothesis-free} analysis is well suited for deriving new insight into the data, such as identifying confounding features or potential targets (e.g. deriving risk factors for breast cancer associated features).
%%
}
%%
%Input formulas reflect \emph{hypotheses} about the data.
\emph{Hypotheses} about the data are reflected using input formulas.
%%
Using the operators, dynamic variables and dataset features, many different assumptions can be expressed.
%%
To support the \emph{hypothesis-free} analysis, we provide a default formula:

$Z \sim X + Y$.
%%
It represents all possible combinations of two independent features w.r.t. all features in the data set, since we do not know which features are of interest.
%%
Each slice represents a different target feature.
%%
It is therefore suitable for an explorative analysis to give a general impression about relationships in the data set.

@@

%%
$Cancer \sim X + Y + Z$ is the formulation of a hypothesis, where the specific feature \textit{Cancer} is analyzed.
%%
All combinations of three independent features with the target are analyzed through this \emph{3D Regression Heat Map}.

$Cancer \sim X + Y + Z + feature_1:feature_2$ encodes more assumptions.
%%
This formula models the hypothesis of an interaction between $feature_1$ and $feature_2$ (denoted with `$:$') being relevant for the target feature, but it is not clear how other feature combinations influence the result.
%%
Therefore, this interaction is incorporated for all $X$, $Y$ and $Z$ values as independent features.
%%

$Cancer \sim X + Y + Z$ subtracted with the regression metric from $Cancer \sim Age$ excludes the confounding effect that age has in view of the target $Cancer$ feature.
%%
This is achieved through \emph{3D Regression Heat Map} comparison.
%%
